---
title: "Final Project - Getting and Cleaning Data"
author: "Marcos"
date: "20/04/2021"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(readr)
library(tidyverse)
```

## Step 1: Loading the data

The content of the .txt files are loaded using the read_table function as follows:

```{r, message=FALSE, warning=FALSE}

# Test data

test_set <- read_table(file = "data/UCI HAR Dataset/test/X_test.txt", col_names = F)
test_label <- read_table(file = "data/UCI HAR Dataset/test/y_test.txt", col_names = F)


# Train data

train_set <- read_table(file = "data/UCI HAR Dataset/train/X_train.txt", col_names = F)
train_label <- read_table(file = "data/UCI HAR Dataset/train/y_train.txt", col_names = F)


# Activities labels

act_labels <- read_table(file = "data/UCI HAR Dataset/activity_labels.txt", col_names = F)

# Features

features <- read_table(file = "data/UCI HAR Dataset/features.txt", col_names = F)


```



## Step 2: Tidying the data

Firstly, the "features" data was cleaned by removing unecessary characters. Secondly, we defined the "features" data as column names for the data frame.Thirdly, we binded the "subject" data to the dataframe. Lastly, we extracted only columns containing the terms "mean" and "std", as it was demanded.

```{r message=FALSE, warning=FALSE}

# Naming columns
features2 <- str_replace(string = features$X1, pattern = "^[0-9]+", replacement = "")
colnames(test_set) <- features2
colnames(train_set) <- features2

#Adding train label

test_set <- bind_cols(test_label, test_set)
train_set <- bind_cols(train_label, train_set)


# Adding subject train

subject_train <- read_table(file = "data/UCI HAR Dataset/train/subject_train.txt", col_names = F)
subject_test <- read_table(file = "data/UCI HAR Dataset/test/subject_test.txt", col_names = F)

data_test <- bind_cols(subject_test, test_set)
data_train <- bind_cols(subject_train, train_set)

names(data_test)[1] <- "subject"
names(data_train)[1] <- "subject"

names(data_test)[2] <- "activity ID"
names(data_train)[2] <- "activity ID"

# Extracting only columns containing mean and std

data_test_selected <- data_test %>%
  select(contains("subject"),
         contains("activity ID"),
         contains("mean"),
         contains("std"),
         -contains("angle"))

data_train_selected <- data_train %>%
  select(contains("subject"),
         contains("activity ID"),
         contains("mean"),
         contains("std"),
         -contains("angle"))


```

## Step 3: Merging both dataframes

In this part, we binded both "test" and "train" dataframes into a single one. Then, we merged the resulting dataset using the "inner_join" function, so that every observation would show its related "activity name". The key used was the "activity ID" number.

```{r}

# Merging both sets

full_data_selected <- bind_rows(data_test_selected, data_train_selected)
full_data_selected_2 <- inner_join(x = act_labels, y = full_data_selected, by = c("X1" = "activity ID"))

names(full_data_selected_2)[1] <- "activity_id"
names(full_data_selected_2)[2] <- "activity"

head(full_data_selected_2)
```


## Step 4: Pivoting the dataframe

Lastly, we transformed the "wide dataframe" into a "long format" using the pivot_longer function. The values of the observations now belong to the "measurement" column and each variable to the "variable" column. 

```{r}

# Pivoting the DF

data_tidy <- full_data_selected_2 %>%
  pivot_longer(cols = -c(activity_id, activity, subject),names_to = "variable", values_to = "measurement") %>%
  select(subject, activity_id, activity, variable, measurement)

head(data_tidy)

```

## Step 5: Calculating the average grouped by activity

To calculate the average of the measurement by activity, we just need to use the group_by function:

```{r, message=FALSE, warning=FALSE}

# average of each variable grouped by activity

average_by_act <- data_tidy %>%
  group_by(activity, variable) %>%
  summarize(average = mean(measurement))

head(average_by_act)

```



## Step 6: Calculating the average grouped by subject

```{r, message=FALSE, warning=FALSE}

# average of each variable grouped by subject

average_by_subject <- data_tidy %>%
  group_by(subject, variable) %>%
  summarize(average = mean(measurement))

head(average_by_subject)

```

