---
title: CODEBOOK"
author: "Marcos"
date: "20/04/2021"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(readr)
library(tidyverse)
```

## Describing the variables

Test data and train data contains the measurement values produced by the Samsung watch. 

Test label and train label contain the type of activity that produced the measurements (walking, laying etc)

```{r, warning=FALSE, message=F}
# Test data

test_set <- read_table(file = "data/UCI HAR Dataset/test/X_test.txt", col_names = F)
test_label <- read_table(file = "data/UCI HAR Dataset/test/y_test.txt", col_names = F)

# Train data

train_set <- read_table(file = "data/UCI HAR Dataset/train/X_train.txt", col_names = F)
train_label <- read_table(file = "data/UCI HAR Dataset/train/y_train.txt", col_names = F)

head(train_set)
head(train_label)

```


The activity label describes the name of each activity studied. It´s used, in this work, to describe the activity ID number.

```{r, warning=FALSE, message=FALSE}

# Activities labels

act_labels <- read_table(file = "data/UCI HAR Dataset/activity_labels.txt", col_names = F)
head(act_labels)

```

Features data contain the variables that were measured by the Samsung watch, as it is described in the website:

"The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions."

```{r, message=FALSE, warning=FALSE}


# Features

features <- read_table(file = "data/UCI HAR Dataset/features.txt", col_names = F)
features2 <- str_replace(string = features$X1, pattern = "^[0-9]+", replacement = "")
head(features2)

```

